---
title: "Graph Theory"
author: "Navankur Verma"
date: "10/06/2020"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 2
    toc_float:
      smooth_scroll: FALSE
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 8)

```

# Graphs using `igraph`

Graph using formula notation:
```{r, message=FALSE}
library(igraph)

g <- graph.formula(1-2, 1-3, 2-3, 2-4, 3-5, 4-5, 4-6, 4-7, 5-6, 6-7)
plot(g)
```

In this case we are explicitly specifying the edges of g, which are all undirected.

```{r}
g2 <- graph.formula(1-+2, 1-+3, 2++3, 1-+4)
g3 <- graph.formula(Sam-+Mary, Sam-+Tom, Mary++Tom)
plot(g2)
plot(g3)
```

Graph using Edge list:

```{r}
el <- matrix(c(1,2, 1,3, 1,4, 2,3, 3,4), byrow = T, ncol = 2)
g4 <- graph_from_edgelist(el)
plot(g4)
```

Graph using Adjacency Matrix:

```{r}
adj <- matrix(round(runif(n = 9)), nrow = 3)
diag(adj) <- 0 #main condition for 'no-self' loop, diag elements are zero.

g5 <- graph_from_adjacency_matrix(adj)
plot(g5)
```


# Printing and MetaData

```{r}
library(sand) #various graphs in 'igraph' format
data(lazega)
lazega <- upgrade_graph(lazega)# needed due to recent updates of igraph package

V(lazega) #Vertices

E(lazega) #Edges

lazega #Summary print
```

In the first line, UN means undirected, 36 is the number of nodes and 115 the number of edges.

Additional metadata attached to the graph are called attributes and can be discovered with `list.???.attributes(lazega)`, where `???` can be one of graph, vertex, edge. We can print out the available details of graphs and even update them:
```{r}
list.vertex.attributes(lazega)

get.vertex.attribute(lazega, name = "Gender")

lazega = set.graph.attribute(lazega, "name", "Lazega Lawyers")
list.graph.attributes(lazega)
lazega
```

Its Adjacency matrix:
```{r}
adj <- get.adjacency(lazega)
str(adj)
```

Easy to work with:
```{r}
adj <- as.matrix(adj)
head(adj)

image(adj)
```

# Plotting

Layouts:
```{r}
layout <- layout.fruchterman.reingold(lazega) # most commonly used layout, also used by default
layout_circle <- layout_in_circle(lazega) # sometimes used

plot(lazega, layout = layout)
plot(lazega, layout = layout_circle)
```


Aesthetics:
```{r}
# igraph.options(
#   vertex.color = V(lazega)$Office,
#   vertex.label = NA,
#   vertex.size = 1.5 * degree(lazega),
#   edge.width = 2,
#   edge.color = "grey50",
#   edge.curved = 0.5
# )

#or set options directly in the plot command:
plot(
  lazega,
  vertex.color = V(lazega)$Office,
  vertex.label = NA,
  vertex.size = 1.12 * degree(lazega),
  edge.width = 2,
  edge.color = "grey50",
  edge.curved = 0
)

shapes <- c("circle","square")

#Office location indicated by color, type of practice indicated by vertex shape, vertex size proportional to years with firm

plot(
  lazega,
  vertex.color = V(lazega)$Office,
  vertex.shape = shapes[V(lazega)$Practice],
  vertex.size = 0.5* V(lazega)$Years
)

# 3 most central nodes based on different centrality measures:

plot(
  lazega,
  vertex.color = (centr_betw(lazega)$res >= sort(centr_betw(lazega)$res, decreasing = T)[3]) ,
  main = "Betweenness Centrality\nBased on number of shortest paths which a nodes lies on"
)

plot(
  lazega,
  vertex.color = (centr_clo(lazega)$res >= sort(centr_clo(lazega)$res, decreasing = T)[3]) ,
  main = "Closeness Centrality\nBased on Avg Geodesic Distant (Shortest Path)"
)

plot(
  lazega,
  vertex.color = (centr_eigen(lazega)$vector >= sort(centr_eigen(lazega)$vector, decreasing = T)[3]) ,
  main = "Eigen Centrality\nBased on important nodes a node connects with"
)

plot(
  lazega,
  vertex.color = (page_rank(lazega)$vector >= sort(page_rank(lazega)$vector, decreasing = T)[3]) ,
  main = "Page Rank\nKatz Centrality with extension of Penalization for higher degree"
)

plot(
  lazega,
  vertex.color = coreness(lazega)+1,
  main = "k-Cores\nNodes which are connected to at least k others nodes"
)
legend("topright", legend = c("0-Core","1-Core","2-Core","3-Core","4-Core","5-Core"), fill = categorical_pal(6))

# plot(
#   lazega,
#   vertex.color = adjustcolor(coreness(lazega)+1, 0.5),
#   main = "k-Cores", sub = "Nodes which are connected to at least k others nodes"
# )
# legend("topright", legend = c("0-Core","1-Core","2-Core","3-Core","4-Core","5-Core"), fill =adjustcolor(unique(coreness(lazega)+1), 0.5))

#removing nodes with based on null degrees:
reduced_lazega <- delete.vertices(lazega, which(degree(lazega) == 0))
plot(reduced_lazega)
```

# Network Summaries

```{r}
data(ppi.CC) #Protien Interactions graph
ppi.CC <- upgrade_graph(ppi.CC)
```

Global clustering coefficient is:
```{r}
transitivity(ppi.CC)
```

The degrees and the degree distribution:

```{r}
d.yeast <- degree(ppi.CC)

hist(d.yeast,col="blue",
     xlab = "Degree", ylab = "Frequency",
     main = "Degree Distribution")

barplot(table(d.yeast), xlab = "Degree", ylab = "Frequency",
        main = "Degree Distribution", col = "SkyBlue")

```

On log-log scales:

```{r}
dd.yeast <- degree.distribution(ppi.CC)
d <- 0:max(d.yeast)
ind <- (dd.yeast != 0)

plot(d[ind], dd.yeast[ind], log = "xy", col = "blue",
     xlab = c("Log-Degree"), ylab = c("Log-Proportion"),
     main = "Log-Log Degree Distribution")
```

Degree of the neighbours of a node:

```{r}
a.nn.deg.yeast <- graph.knn(ppi.CC,V(ppi.CC))$knn
plot(d.yeast, a.nn.deg.yeast, log="xy",
     col = "goldenrod", xlab = c("Log Vertex Degree"),
     ylab = c("Log Average Neighbour Degree"))
```

Vertex (node) centrality:
```{r}
ppi_cent_1 <- centr_eigen(ppi.CC)$vector # eigenvector centrality
ppi_cent_2 <- page_rank(ppi.CC)$vector # Page-Rank centrality
ppi_cent_3 <- centr_clo(ppi.CC)$res # closeness centrality
ppi_cent_4 <- centr_betw(ppi.CC)$res # betweenness centrality
chosen_centrality <- ppi_cent_1 # change as needed
V(ppi.CC)[order(chosen_centrality, decreasing = T)[1:5]]
```

Edge Centrality:
```{r}
ppi_cent_5 <- edge.betweenness(ppi.CC)
E(ppi.CC)[order(ppi_cent_5, decreasing=T)[1:3]]
```

```{r}
cores <- coreness(ppi.CC) #nodes which are connected to at least k other 
```
